A Constraint Solving Approach to Parikh Images of Regular Languages

### 2024-02-19 evaluation.tex
Update results from SMT-COMP

\begin{table}
  \begin{center}
  \input{experiments/qf_slia_comparison.tex}
  \end{center}
  \caption{Number of solved benchmarks in the set of quantifier-free string
    linear integer arithmetic (QF\_SLIA) at SMT-COMP 2023. The numbers are from
    the competition results, except for CA-Str, which is executed by us (600s
    timeout, AMD Opteron 2220 SE), and \Ostrich+CA, a virtual portfolio. }
  \label{tab:solve-status-smt-comp}
\end{table}

@@

For our experiments, we modified the CE-Str solver to apply \Catra{}
instead of the previous baseline method, resulting in a new back-end
\textbf{CA-Str}. We created also a new virtual portfolio~\Ostrich{}+CA
by combining the three solvers in \Ostrich{}~1.3 with \textbf{CA-Str}.

@@

We evaluate the different solvers in \cref{tab:solve-status-smt-comp}
on 21\,938~benchmarks from the single-query QF\_SLIA track at SMT-COMP 2023, as
QF\_SLIA was the most challenging track for \Ostrich{} in the
competition.
%
As can be seen from the table, integrating \Catra{} as a back-end
leads to gains both on satisfiable and unsatisfiable problems. On
satisfiable benchmarks, the combination \Ostrich+CA is still
outperformed by cvc5 and z3alpha. On unsatisfiable benchmarks,
\Ostrich+CA narrowly beats the other solvers, which is promising given
that also \Ostrich{}~1.3, cvc5, and z3alpha show very strong
performance on this class of benchmarks.

The results show that \Calculus{} can be used to enhance the performance of an
automaton-based string solver. The results should be considered preliminary,
however, as we believe that a deeper integration of \Catra{} into string solvers
can lead to significant performance gains.

@@

The most obvious threat to validity would be an unsound implementation. To
address this we have validated all reported solutions made by \Calculus{} with
\Nuxmv{}. A previous version contained a race with random restarts during
product materialisation causing non-deterministic unsoundness in \numprint{0.7}
of instances. Since addressing that bug we have observed no further soundness
issues.

>>>>

\begin{table}
  \begin{center}
  \input{experiments/qf_slia_comparison.tex}
  \end{center}
  \caption{Number of solved benchmarks in the set of quantifier-free strings with
    linear integer arithmetic constraints (QF\_SLIA) at SMT-COMP~2023. The numbers are from
    the competition results, except for CA-Str, which is executed by us on the same 
    cluster as the competition with the same resources, and the two virtual portfolio solvers
    \Ostrich+CA and \textsc{Competition} which aggregate the best results from \Ostrich{}/CA-Str and 
    all the competition results for non-\Ostrich{} solvers respectively. }
  \label{tab:solve-status-smt-comp}
\end{table}

@@

For our experiments, we modified the CE-Str solver to apply \Catra{}
instead of the previous baseline method, resulting in a new back-end
\textbf{CA-Str}. Combining the results from SMT-COMP for \Ostrich{}~1.3 
with our new results running CA-Str we construct a virtual 
portfolio~\Ostrich{}+CA that simulates running CA-Str as a fourth
back-end to \Ostrich{}.

@@

We extend the results of SMT-COMP 2023 with our modified \Ostrich
on the single-query string solving with linear integer arithmetic constraint
track (QF\_SLIA) in~\cref{tab:solve-status-smt-comp}. We obtain
the results by executing \Ostrich~1.3 with CA-Str using
the same benchmarking infrastructure (the StarExec cluster)
and configuration that ran SMT-COMP 2023, combining our new results for CA-Str with the
published results of SMT-COMP 2023. Where avaiable, we use the revised, out-of
competition version of the results for solvers with bugs.

We picked this track since
\Ostrich{} already performed well on the other two string solving tracks. In fact,
every solver, including \Ostrich{}, handled every benchmark in QF\_SNIA within the timeout.
%
As can be seen from \cref{tab:solve-status-smt-comp}, integrating \Catra{} as a back-end
leads to gains both on satisfiable and unsatisfiable problems. On
satisfiable problems, the combination \Ostrich+CA is still
outperformed by cvc5 and z3alpha. On unsatisfiable benchmarks,
\Ostrich+CA narrowly beats the other solvers,
which is promising given that \Ostrich{}~1.3 (without CA-Str), cvc5, and z3alpha show very strong
performance on this class of benchmarks. Similar improvements
occur in QF\_S, they are much smaller (an additional 2 unsatisfiable, 4 satisfiable 
solved out of~8\,847 in total).

The results show that \Calculus{} can be used to enhance the performance of an
automaton-based string solver. The results should be considered preliminary,
however, as we believe that a deeper integration of \Catra{} into string solvers
can lead to significant performance gains. In particular, the integration layer 
is currently too shallow to allow Ostrich to learn clauses generated by \Catra{}
and additionally incurs subprocess-spawning overhead.

@@

The most obvious threat to validity would be an unsound implementation. To
address this we have validated all reported solutions made by \Calculus{} with
\Nuxmv{}. A previous version contained a race with random restarts during
product materialisation causing non-deterministic unsoundness in \numprint{0.7}
of instances. Since addressing that bug we have observed no further soundness
issues, with the exception of one erroneous \textsc{Unsat} response on the QF\_S track of 
SMT-COMP (\texttt{QF\_S\/20230329-automatark-lu\/instance08425.smt2}) that likely occurred
in the interaction surface between \Ostrich{} and \Catra{}.
